---
title: "Objective 1"
---

## Probability as a foundation of statistical modeling

Course Objective:

> Describe probability as a foundation of statistical modeling, including inference and maximum likelihood estimation

Probability is a fundamental concept in statistical modeling, providing a quantitative measure of the likelihood of an event occurring within a given set of possible outcomes. It plays a vital roles in various aspects of statistics, including inference and maximum likelihood estimation.

Some basic examples to illustrate probability distributions, statistical modeling, inference, and maximum likelihood estimations includes:

### 1. Probabitlity Distributions

```{r probability-distributions}
# Uniform distribution
min <- 0 # lower limit
max <- 1 # upper limit

# Generate 1000 random numbers from a uniform distribution
x <- runif(1000, min = min, max = max)
```

```{r probability-distributions-2}
set.seed(631)

# Generate a sample dataset.
x <- seq(1, 100)
# True relationship: y = 2x + noise
y <- 2 * x + rnorm(100, mean = 0, sd = 10)

# Fit linear regression model
model <- lm(y ~ x)

# Model summary
summary(model)
```

### 2. Statistical Modeling - Linear Regression

```{r statistical-modeling}
set.seed(42)

# Generate sample dataset
x <- seq(1, 100)
# true relationship: y = 2x + noise
y <- 2 * x + rnorm(100, mean = 0, sd = 10)

# Fit linear regression model
model <- lm(y ~ x)

# Model summary
summary(model)
```

### 3. Inference - Confidence Interval and Hypothesis Testing:

```{r}
# Confidence interval
conf_int <- confint(model, level = 0.95)  # 95% confidence interval for model parameters
print(conf_int)

# Hypothesis testing (t-test)
t_test <- summary(model)$coefficients  # t-test results for model coefficients
print(t_test)
```

### 4. Maximum Likelihood Estimation - Logistic Regression:

```{r}
# Simulate data
set.seed(42)
x <- rnorm(100)
z <- 1 + 2 * x
prob <- 1 / (1 + exp(-z))
y <- rbinom(100, size = 1, prob = prob)  # binary response variable

# Fit logistic regression model using maximum likelihood estimation
model <- glm(y ~ x, family = binomial(link = "logit"))

# Model summary
summary(model)
```

The above example demonstrate working with probability distributions, fitting statistical models, perform inference and apply maximum likelihood estimation in the context of linear and logistic regression.
